---
title: "Analyzing Nickelback's Lyrics"
author: "Matt M"
subtitle: "A Web-Scraping and Text Analysis Tutorial"
output: html_notebook
---

## Background
Nickelback's music is arguably one of the worst exports that Canada has ever produced. It's rumored that Nickelback was created by God as payback for the sins of mankind. In a ranking by Gitmo detainees, Nickelback was ranked as the third-worst torture device (Acid drips on bare skin was #1 and electrocution was #2). But in all seriousness, my colleagues and I were curious about how Nickelback is actually a commercially-viable band. With all the recycled content and rhyming for the sake of rhyming, you'd think people would have just stopped listening.  

To better understand Nickelback, I've decided to take one for the team and scrape their lyrics and perform some text analysis.  

## Setting-Up for Scraping and Text Analysis
### Special functions
As always, special functions are specified first. `wipe_html` is handy for eliminating annoying HTML tags. It essentially takes a HTML string, cleans out the HTML tags, and returns the non-HTML contents of that string. It'll come in handy in any web-scraping you do, so I highly recommend you hang onto it.
```{r}
# function to scrub html out of text ----
wipe_html <- function(str_html) {
        gsub("<.*?>", "", str_html)
}
```

### Libraries
I'll be using Hadley Wickham's `dplyr` and `rvest`. `dplyr` is the gold standard for R data wrangling, and `rvest` is just convenient for scraping web data.
```{r}
# libraries ----
library(rvest)
library(dplyr)
```

### Web-Scraping
#### Get the song list
Finding a song list that wasn't totally jacked up took a little searching. I settled on MetroLyrics as the songs were cleanly listed and duplicates weren't found. I can't vouch that these are all of Nickelback's songs, but what do you care? It's Nickelback.  

This code chunk scrapes the table with the song titles, selects the name of the song and the year it was released, and creates a column called `url_name` that I'll need in the next code chunk.  
```{r}
# get the HTML code from metrolyrics ----
songs_url <- "http://www.metrolyrics.com/nickelback-lyrics.html"
html_code <- read_html(songs_url)

# get songs from song title HTML table ----
html_songs <- html_node(html_code, "#popular > div > table")
songs_df <- html_table(html_songs)

# drop name column to lowercase and remove columns ----
names(songs_df) <- tolower(names(songs_df))
songs_df <- songs_df %>% select(name, year)

# clean up the songs_df "song_title" variable ----
songs_df_clean <- songs_df %>% 
        mutate(url_name = gsub("\\'|\\.|&", "", name),
               url_name = gsub("\\(|\\)", " ", url_name),
               url_name = tolower(url_name),
               url_name = gsub(" ", "-", url_name),
               url_name = gsub("--", "-", url_name))
```

### Getting the lyrics
Looping over the `url_name` variable from the previous code chunk made the most sense for this.  

In this code chunk, I'll create two empty vectors for the lyrics and album names. My `for`-loop will then loop over each element in `url_name`, paste it into a hyperlink, scrape the lyrics and album name, scrub out the HTML, and place each captured piece into their respective vectors. The data frame `album_lyrics` will hold the contents of both vectors.
```{r}
# navigate to each song's URL and scrape the album name and lyrics ----
## empty vectors
lyrics <- c()
album <- c()
## specify row number to add to data frame
number <- 1
## for-loop and create data frame from the two vectors
for(i in seq_along(songs_df_clean$url_name)) {
        for_url_name <- songs_df_clean$url_name[i]
        ## filter for row with song title
        sub_row <- filter(songs_df_clean, url_name == for_url_name)
        ## create url
        paste_url <- paste0("http://www.metrolyrics.com/", for_url_name,
                            "-nickelback.html")
        ## open connection to url
        for_html_code <- read_html(paste_url)
        ## scrape lyrics via selector path
        for_lyrics <- html_node(for_html_code, "#lyrics-body-text")
        ## scrub html and control characters out of the lyrics
        for_lyrics <- wipe_html(for_lyrics)
        for_lyrics <- gsub("[[:cntrl:]]", " ", for_lyrics)
        ## scrape album name
        for_album <- html_node(for_html_code, "#album-name-link")
        ## scrub html out of album name
        for_album <- wipe_html(for_album)
        ## add for_lyrics and for_album to respective vectors 
        lyrics[number] <- for_lyrics
        album[number] <- for_album
        number <- number + 1
        ## optional: add in 5 second delay to avoid IP block
        #Sys.sleep(5)
}
album_lyrics <- data.frame(album = album, lyrics = lyrics)

# bind data frames together ----
if(nrow(songs_df_clean) != nrow(album_lyrics)) {
        stop("data frames have different number of rows")
} else {
        nb_data <- bind_cols(songs_df_clean, album_lyrics)
}
```

### Look at the Data
There are 75 rows in this data frame, but I'll show the first 10 only.
```{r}
# show the first 10 rows of nb_data ----
nb_data[1:10, ]
```

#### Up Next: Text analysis of the lyrics
When I get some more time this week or next, I'll walk through how to preprocess the lyrics and do some basic text analytics.
